---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

I chose to do all processing within the section that was relevant to the processing needed. I think this makes it easier to understand why I did what I did when I did it :)

```{r echo=TRUE}
activity <- read.csv('activity.csv')

```
## What is mean total number of steps taken per day?

```{r echo=TRUE}
tot_Steps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
hist(tot_Steps, breaks=10, col='light blue', main = 'Total Number of Steps Per Day', xlab = 'Total Steps')
summary(tot_Steps)
```
Using the summary function, we can see that the mean number of steps taken per day is 9350 and the median is 10400


## What is the average daily activity pattern?
```{r echo=TRUE}
Avg_Steps <- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)
plot(Avg_Steps ~ names(Avg_Steps), type='l', col='dark blue', main = 'Average Steps per 5 Minute Time Interval', xlab = 'Time Interval', ylab = 'Average Steps')
max(Avg_Steps)
which.max(Avg_Steps)

```
We can see, using the max and which.max functions, that the interval with the highest average number of steps is the 835 interval, and the max value is 206.2

## Imputing missing values

```{r echo=TRUE}
activity_imputed <- activity
activity_imputed$steps[is.na(activity_imputed$steps) == TRUE] = mean(activity_imputed$steps, na.rm=TRUE)
tot_StepsImputed <- tapply(activity_imputed$steps, activity_imputed$date, sum)
 hist(tot_StepsImputed, breaks=10, col='light blue', main = 'Total Number of Steps Per Day Imputed', xlab = 'Total Steps')
summary(tot_StepsImputed)
summary(tot_Steps)
```
After using the mean value for our missing data, we see our mean and median for total number of steps in a day rise to 10800.


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
activity_imputed$date <- as.POSIXlt(activity_imputed$date)
weekday <- weekdays(activity_imputed$date, abbreviate = TRUE)
activity_imputed$weekday <- 1
activity_imputed$weekday[weekday %in% c('Mon', 'Tue', 'Wed', 'Thu', 'Fri')] = 'weekday'
activity_imputed$weekday[weekday %in% c('Sat', 'Sun')] = 'weekend'
activity_imputed$weekday <- as.factor(activity_imputed$weekday)

weekdayActivity <- subset(activity_imputed, activity_imputed$weekday == 'weekday')
weekendActivity <- subset(activity_imputed, activity_imputed$weekday == 'weekend')
avgWeekdaySteps <- tapply(weekdayActivity$steps, weekdayActivity$interval, mean)
avgWeekendSteps <- tapply(weekendActivity$steps, weekendActivity$interval, mean)

par(mfcol = c(2,1))
plot(avgWeekendSteps ~ names(avgWeekendSteps), type = 'l', xlab = 'Weekend Interval', ylab = 'steps')
plot(avgWeekdaySteps ~ names(avgWeekdaySteps), type = 'l', xlab = 'Weekday Interval', ylab = 'steps')

```

We can see from our plots that there seems to be more total activity during the weekend, and the number of steps seem to be more evenly distributed throughout the day than they are during the weekdays.